Stack Safety is a Security Property

### 2021-04-13 tex/main.tex
INtro

\bcp{We should mention stack data and coroutines, I think?  Maybe in the
  context of an expectation-lowering hedge, like:}%
For the sake of simplicity, most of our presentation assumes a RISC-V
machine architecture with a basic subroutine calling convention, where
parameters are passed in registers.  We also show how stack-allocated
parameter passing and coroutines can be accommodated.

@@

The call stack is a perennial target for low-level attacks
based on memory-safety vulnerabilities, leading to consequences
that range from leakage or corruption of private stack data to
control-flow hijacking. To detect or prevent such attacks, a profusion of
software and hardware protections have been proposed,
%
including stack canaries~\citep{Cowan+98},
bounds checking~\citep{NagarakatteZMZ09,NagarakatteZMZ10,DeviettiBMZ08},
split stacks~\citep{Kuznetsov+14},
shadow stacks~\citep{Dang+15,Shanbhogue+19},
capabilities~\citep{Woodruff+14,Chisnall+15,Skorstengaard+19,Skorstengaard+19b,Tsampas+19,Georges+21},
and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}. \ifaftersubmission\apt{Mostly from
  nick; there could be more}\bcp{Yes, going back to MIT days---we should
  include several more of these, if only to give readers the impression that
this is a well-studied mechanism (so formalizing its protections is
useful).}
\fi
%
The protections offered by such mechanisms are generally described in terms
of a set of concrete examples of attacks that they can prevent, such as corruption
of return addresses, buffer overflows, or use of uninitialized variables. This leaves the
task of formulating a more general characterization to the reader's intuition.
Moreover, the details of these mechanisms can be quite intricate
and often involve trading off protection and efficiency.
Thus, it would be useful to have a precise, generic, and formal
specification for stack
safety, both as a basis for comparing the security claims of different
enforcement techniques and for validating more rigorously that these claims
are met by particular implementations.

%\bcp{Does use of unitialized variables lead to stack safety attacks?\apt{sure, via buffer overflow}}
%with the notion usually being
%defined negatively: through examples of stack-based exploits
%\rb{usually though not always; as noted below there are other
%  approaches}\bcp{Yes, we have to be very careful about this}.
%\leo{Examples at a high level}: such as -- does not overwrite return addresses, or ...
The most abstract characterization of which we are aware is in
recent work by \citet{Skorstengaard+19}, which defines stack safety as the conjunction of two
properties:
{\em local state encapsulation (LSE)} and {\em well-bracketed control flow (WBCF)}.
Informally, LSE says that the contents of
a caller's stack frame are not read and written while its callees (and their callees, etc.) are
executing, and WBCF says that callees always
return to the point just after their call (or don't return at all).
\citet{Skorstengaard+19} formalize these properties by defining an idealized machine
in which they visibly hold by construction; an enforcement mechanism (in their case, hardware
capabilities)
can then be shown to provide stack safety by giving a fully abstract embedding function
from the idealized machine to an ordinary machine equipped with the enforcement.
%%
%% While this work represents an important step in the right direction, we believe
%% there is still room to craft an operational characterization
%% of stack safety. \apt{Perhaps our rhetoric should be more micro-policy-specific here?}
%% In other words, we seek formal criteria that can be applied
%% to a mechanism for protecting stack-based low-level programs to judge
%% whether or not it actually ``guarantees stack safety.''
%% \rb{The remainder of this paragraph may be diving too quickly, would a reader be
%% confused by the references to these types of properties, yet to be introduced?}
%% We begin with a strong
%% local state encapsulation property that closely models the behavior of eager
%% micro-policies from \citet{DBLP:conf/sp/RoesslerD18}. This ``eager'' property
%% is too strong to characterize stack safety; reasonable enforcement mechanisms
%% exist that violate it. One such mechanism is a ``lazy'' tag policy, similar to
%% the one in \citet{DBLP:conf/sp/RoesslerD18}. We weaken local state encapsulation
%% to support this technique and argue that the new property, defined relative to an
%% arbitrary notion of observation. This property is still strong enough to prevent
%% the bad behaviors we associate with unsafe stack behavior, but weak enough that
%% other reasonable enforcement mechanisms can enforce it.

In this work, we take the (informal) definitions of WBCF\apt{or maybe not?}  and LSE as a starting
point, but choose a different approach to formalizing them, based on the
technical framework of language-based security~\citep{??}. We define
trace-based properties of \emph{stepwise stack integrity} and
\emph{lockstep stack confidentiality}.\apt{Could we just say ``stepwise'' for both?}
In applying these security concepts, we treat each caller as needing protection
from its callee, even as that callee may need protection from its callee in turn.
This differs from common practice in language-based security in that the
confidentiality property applies to many nested subtraces within a single execution.
\sna{How novel is this nesting?} \apt{Lost the reasoning principle that was here?}
{\em Integrity} says that a caller's local data is protected from modification by its callee,
and its {\em stepwise} interpretation formalizes this very naturally: a caller's local data
is never modified during a call. {\em Confidentiality} says that a caller's data is protected
from being leaked into its callee, and also that the callee cannot see uninitialized data.
As a reasoning principle, confidentiality guarantees that the callee's
behavior is determined only by a predictable interface that does not include the caller's
local state or the uninitialized portion of the stack. This is formalized as a
{\em noninterference}-style property that, rather than comparing pairs of whole program traces,
compares a pair of subtraces for each call, nested within other calls. The term
{\em lockstep} describes the behavior of these pairs: they step together, and any
deviation is recognized as a violation of confidentiality.
Together, these properties capture the rough intuition of LSE, but are somewhat stronger:
confidentiality contains temporal protections that LSE does not address.\apt{Maybe want to
  say a little more about this here, since it is a novelty wrt/ existing work?}

@@

Part of the challenge in fully\apt{omit?} characterizing stack safety is that in a low-level machine
without built-in stacks, we are at the mercy of architectural
choices that do not relate to the abstract notion of stack safety\apt{a bit vague. what kind of choices?}.
We abstract over these choices
as much as possible, so that our model can be applied to a wide range of architectures
and enforcement mechanisms.

\apt{Intra-paragraph flow seemed awkward.}
A second challenge is
% the diversity of enforcement mechanisms APT:already implied above
enforcement mechanisms may weaken their protections that in the name of efficiency,
while arguing that they still capture stack safety in a meaningful sense.
\apt{Remainder of this paragraph might be better \emph{after} the R+DH para below.}
This calls \apt{not
  immediately obvious why. is this true for anything other than laziness?}
for an {\em extensional}
characterization of stack safety that describes the externally observable behavior
of a stack-safe system. In contract, our stepwise and lockstep properties are {\em intensional},
focusing on the internal behavior.

@@

To illustrate the applicability of these properties, we perform randomized testing to
validate that \citeauthor{DBLP:conf/sp/RoesslerD18}'s Depth Isolation micro-policy enforces
them. Randomized testing is a relatively low-cost way to validate a potential enforcement
mechanism; it finds flaws in the enforcement fast enough to be integrated into the workflow
of enforcement development \citep{TestingNI:ICFP}. To demonstrate that the testing adequately
covers potentially erroneous programs, we test intentionally buggy policies alongside
the real one. If the testing can differentiate correct implementations from buggy ones,
it is an argument in favor of their correctness.
%\apt{more to say?}\bcp{Yes---partly to underscore our claim that this is a
%  significant contribution, and partly to give the reader some context for
%  why this is part of our story.  E.g., ``One nice thing we can do with our
%  spec is to use it to evaluate the claims of existing or future proposals
%  for enforcement mechanisms.  One relatively inexpensive way to carry out
%  such an evaluation is to see whether property-based testing against this
%  spec can distinguish correct implementations of the mechanism from buggy
%  ones...''}
Of course, as our testing reveals, \citeauthor{DBLP:conf/sp/RoesslerD18}'s Lazy Tagging and
Clearing micro-policy does not enforce stepwise integrity.
Unfortunately the existing Lazy Tagging and Clearing
micro-policy does not quite enforce observable confidentiality either. %, as we show in
%testing~\apt{really??},
But formalizing the ideal property allows us to see how it falls short -- a specific
violation of the temporal aspect of confidentiality.
We propose  a variant of Lazy Tagging and Clearing that we believe \emph{does} enforce
observational integrity, at some performance cost.\apt{but we're not testing that, right?}
\rb{Not sure about the property part right now, the machine side of this is
  pretty much done}
\sna{It's aspirational so far, but there's no fundamental reason we can't test it. We may
  need a caveat about how it's harder to test?}\apt{As per email discussion,
  maybe we could have a property variant that \emph{does} match depth-based Lazy Tagging and Clearing,
  by loosening the temporal safety properties?}

Our initial description of these properties assumes a single simple stack with no sharing between callers and
callees: all parameters and return values are passed in registers.  However,
we can refine it to allow passing of scalar stack data, granting a callee
access to the caller's data. (We do not model more sophisticated sharing, because to do
so realistically would require a separate protection scheme, such as a capability
model.\bcp{This is all a bit mysterious} While this is quite feasible in a tag-based enforcement mechanism, it is properly
a separate policy outside the scope of stack safety. A tag-based approach can easily
compose multiple such policies.)
\rb{So no pointer passing? Pass by reference is now discussed at the end of
  Section 3, however it looks like things get tricky rather quickly.}
\sna{The trick is that if a programmer is passing pointers, we can't presume their
  intent -- how long should we permit memory to be shared? But if the compiler is passing
  pointers, as in pass by reference, we have a clear sense of where sharing should stop.
  \apt{N.B.Doing pass-by-reference properly requires that the passed variable is accessible
    by all descendents, not just immediate callee, which is doable with depth tags but not
    obviously with per-activation tags }}

Finally, we define integrity and confidentiality properties for a coroutine system featuring
finitely-bounded stacks.
Though simplistic, this extension is a step toward concurrency and demonstrates that
our approach is flexible enough to handle more sophisticated styles of
control flow.

>>>>

\bcp{We should mention stack data and coroutines, I think?  Maybe in the
  context of an expectation-lowering hedge, like:}%
For the sake of concreteness, most of our presentation assumes a RISC-V
machine architecture with a basic subroutine calling convention, where
parameters are passed in registers.  We also show how stack-allocated
parameter passing and coroutines can be accommodated.  \bcp{this is also a
  good place to admit that our definitions are tuned for micropolicies, if
  we can find a succinct way to say it.}

@@

The call stack is a perennial target for low-level attacks
based on memory-safety vulnerabilities, leading to consequences
that range from leakage or corruption of private stack data to
control-flow hijacking. To deter or detect such attacks, a profusion of
software and hardware protections have been proposed,
%
including stack canaries~\citep{Cowan+98},
bounds checking~\citep{NagarakatteZMZ09,NagarakatteZMZ10,DeviettiBMZ08},
split stacks~\citep{Kuznetsov+14},
shadow stacks~\citep{Dang+15,Shanbhogue+19},
capabilities~\citep{Woodruff+14,Chisnall+15,Skorstengaard+19,Skorstengaard+19b,Tsampas+19,Georges+21},
and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}. \ifaftersubmission\apt{Mostly from
  nick; there could be more}\bcp{Yes, going back to MIT days---we should
  include several more of these, if only to give readers the impression that
this is a well-studied mechanism (so formalizing its protections is
useful).}
\fi
%
The protections offered by such mechanisms are commonly described in terms
of concrete examples of attacks that they can prevent, such as corruption
of return addresses, buffer overflows, or use of uninitialized variables,
leaving a more abstract characterization to the reader's intuition.
But these mechanisms can be quite intricate
and often require trading off protection and efficiency.
Thus, it would be useful to have a precise, generic, and formal
specification for stack
safety, both as a basis for comparing the security claims of different
enforcement techniques and for validating more rigorously that these claims
are met by particular implementations.

%\bcp{Does use of unitialized variables lead to stack safety attacks?\apt{sure, via buffer overflow}}
%with the notion usually being
%defined negatively: through examples of stack-based exploits
%\rb{usually though not always; as noted below there are other
%  approaches}\bcp{Yes, we have to be very careful about this}.
%\leo{Examples at a high level}: such as -- does not overwrite return addresses, or ...
The most abstract characterization stack safety of which we are aware is in
recent work by \citet{Skorstengaard+19}, which defines stack safety as the conjunction of two
properties:
{\em local state encapsulation (LSE)} and {\em well-bracketed control flow (WBCF)}.
Informally, LSE says that the contents of
a caller's stack frame are not read and written while its callees (and their callees, etc.) are
executing, and WBCF says that callees always
return to the point just after their call (if they return at all).
\citet{Skorstengaard+19} formalize these properties by defining an idealized
machine in which they hold {\em by construction}; an enforcement mechanism (in their case, hardware
capabilities)
can then be shown to guarantee stack safety by exhibiting a fully abstract
embedding function
from the idealized machine to an ordinary machine equipped with this
enforcement mechanism.  \bcp{Many OOPSLA readers/reviewers will not
  understand that sentence.}
%%
%% While this work represents an important step in the right direction, we believe
%% there is still room to craft an operational characterization
%% of stack safety. \apt{Perhaps our rhetoric should be more micro-policy-specific here?}
%% In other words, we seek formal criteria that can be applied
%% to a mechanism for protecting stack-based low-level programs to judge
%% whether or not it actually ``guarantees stack safety.''
%% \rb{The remainder of this paragraph may be diving too quickly, would a reader be
%% confused by the references to these types of properties, yet to be introduced?}
%% We begin with a strong
%% local state encapsulation property that closely models the behavior of eager
%% micro-policies from \citet{DBLP:conf/sp/RoesslerD18}. This ``eager'' property
%% is too strong to characterize stack safety; reasonable enforcement mechanisms
%% exist that violate it. One such mechanism is a ``lazy'' tag policy, similar to
%% the one in \citet{DBLP:conf/sp/RoesslerD18}. We weaken local state encapsulation
%% to support this technique and argue that the new property, defined relative to an
%% arbitrary notion of observation. This property is still strong enough to prevent
%% the bad behaviors we associate with unsafe stack behavior, but weak enough that
%% other reasonable enforcement mechanisms can enforce it.

In this work, we take the (informal) definitions of WBCF\apt{or maybe
  not?}\bcp{Definitely not, if we're following the approach I proposed.
  WBCF should be discussed second, and we should say that, while it is a
  very reasonable property to want, we do not consider
it as part of stack safety (an enforcement mechanism could be intuitively
``safe'' in the information-flow sense while not enforcing it), but rather
as a separate condition that can be stated straightforwardly using the same
technical machinery.}  and LSE as a starting
point, but choose a different approach to formalizing them, based on the
technical framework of language-based security~\citep{??}. We define
trace-based properties of \emph{stepwise stack integrity} and
\emph{lockstep stack confidentiality}.\apt{Could we just say ``stepwise''
  for both?}\bcp{+1}
In applying these security concepts, we treat each caller as needing protection
from its callee, even as that callee may need protection from its callee in turn.
This differs from common practice in language-based security in that the
confidentiality property applies to many nested subtraces within a single execution.
\sna{How novel is this nesting?} \apt{Lost the reasoning principle that was here?}
{\em Integrity} says that a caller's local data is protected from modification by its callee,
and its {\em stepwise} interpretation formalizes this very naturally: a caller's local data
is never modified during a call. {\em Confidentiality} says that a caller's data is protected
from being leaked into its callee, and also that the callee cannot see uninitialized data.
As a reasoning principle, confidentiality guarantees that the callee's
behavior is determined only by a predictable interface that does not include the caller's
local state or the uninitialized portion of the stack. This is formalized as a
{\em noninterference}-style property that, rather than comparing pairs of whole program traces,
compares a pair of subtraces for each call, nested within other calls. The term
{\em lockstep} describes the behavior of these pairs: they step together, and any
deviation is recognized as a violation of confidentiality.
Together, these properties capture roughly the same intuition as LSE, but are somewhat stronger:
confidentiality contains temporal protections that LSE does not address.\apt{Maybe want to
  say a little more about this here, since it is a novelty wrt/ existing work?}\bcp{+1}

@@

One challenge in characterizing stack safety formally is that in a low-level machine
without built-in stacks, we are at the mercy of architectural
choices that do not relate to the abstract notion of stack safety\apt{a bit vague. what kind of choices?}.
We abstract over these choices
as much as possible, so that our model can be applied to a wide range of architectures
and enforcement mechanisms.  \bcp{As APT has worried, this kind of claims of
generality may be getting us onto thin ice.  Maybe remove the second
sentence (and expand the first)?  }

A second challenge is that
% the diversity of enforcement mechanisms APT:already implied above
enforcement mechanisms may weaken their protections that in the name of efficiency,
while arguing that they still capture stack safety in a meaningful sense.
\apt{Remainder of this paragraph might be better \emph{after} the R+DH para below.}\bcp{+1}
This calls \apt{not
  immediately obvious why. is this true for anything other than
  laziness?}\bcp{Agreed.  If we move this later, we can just say that we can
capture the more performant policies with a more extensional variant of the
criteria that is a bit more complex but usefully weaker.  }
for an {\em extensional}
characterization of stack safety that describes the externally observable behavior
of a stack-safe system. In contract, our stepwise and lockstep properties are {\em intensional},
focusing on the internal behavior.

@@

\bcp{This paragraph doesn't apply to the lazy properties, so maybe it should
be moved before they are introduced?}%
To illustrate the applicability of these properties, we perform randomized testing to
validate that \citeauthor{DBLP:conf/sp/RoesslerD18}'s Depth Isolation micro-policy enforces
them. Randomized testing is a relatively low-cost way to validate a potential enforcement
mechanism; it finds flaws in the enforcement fast enough to be integrated into the workflow
of enforcement development \citep{TestingNI:ICFP}. To demonstrate that the testing adequately
covers potentially erroneous programs, we test intentionally buggy policies alongside
the real one. If the testing can differentiate correct implementations from buggy ones,
it is an argument in favor of their correctness. \bcp{Hmm--don't understand
  that sentence}
%\apt{more to say?}\bcp{Yes---partly to underscore our claim that this is a
%  significant contribution, and partly to give the reader some context for
%  why this is part of our story.  E.g., ``One nice thing we can do with our
%  spec is to use it to evaluate the claims of existing or future proposals
%  for enforcement mechanisms.  One relatively inexpensive way to carry out
%  such an evaluation is to see whether property-based testing against this
%  spec can distinguish correct implementations of the mechanism from buggy
%  ones...''}
Of course, as our testing reveals, \citeauthor{DBLP:conf/sp/RoesslerD18}'s Lazy Tagging and
Clearing micro-policy does not enforce stepwise integrity.
Unfortunately the existing Lazy Tagging and Clearing
micro-policy does not quite enforce observable confidentiality either. %, as we show in
%testing~\apt{really??},
But formalizing the ideal property allows us to see how it falls short -- a specific
violation of the temporal aspect of confidentiality.
We propose  a variant of Lazy Tagging and Clearing that we believe \emph{does} enforce
observational integrity, at some performance cost.\apt{but we're not testing that, right?}
\rb{Not sure about the property part right now, the machine side of this is
  pretty much done}
\sna{It's aspirational so far, but there's no fundamental reason we can't test it. We may
  need a caveat about how it's harder to test?}\apt{As per email discussion,
  maybe we could have a property variant that \emph{does} match depth-based Lazy Tagging and Clearing,
  by loosening the temporal safety properties?}

Our initial description of these properties assumes a single simple stack with no sharing between callers and
callees: all parameters and return values are passed in registers.  Later,
we show how to refine it to allow passing of scalar stack data, granting a
callee
access to the caller's data. (We do not model more sophisticated sharing, because to do
so realistically would require a separate protection scheme, such as a capability
model.\bcp{This is all a bit mysterious} While this is quite feasible in a tag-based enforcement mechanism, it is properly
a separate policy outside the scope of stack safety. A tag-based approach can easily
compose multiple such policies.)
\rb{So no pointer passing? Pass by reference is now discussed at the end of
  Section 3, however it looks like things get tricky rather quickly.}
\sna{The trick is that if a programmer is passing pointers, we can't presume their
  intent -- how long should we permit memory to be shared? But if the compiler is passing
  pointers, as in pass by reference, we have a clear sense of where sharing should stop.
  \apt{N.B.Doing pass-by-reference properly requires that the passed variable is accessible
    by all descendents, not just immediate callee, which is doable with depth tags but not
    obviously with per-activation tags }}

Finally, we define integrity and confidentiality properties for a coroutine system featuring
finitely-bounded\bcp{What's that?} stacks.
Though simplistic, this extension is a step toward concurrency and demonstrates that
our approach is flexible enough to handle more sophisticated styles of
control flow.
